---
title: "Elected Representatives - Analysis"
author: "Melvin"
date: "2023-08-04"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 5
  pdf_document:
    toc: yes
    toc_depth: '5'
  editor_options:
    chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(tidy = T)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(fig.height = 6)
knitr::opts_chunk$set(fig.width =  8)

source("base_cm_new.R")
source("base_mp.R")
```

## CM Analysis

### Information about the Dataset

__List of states in the dataset__

```{r}
df_party %>% select(-c(date_between,inc_states:total_valid_states)) %>% names()
```

__Years covered in the dataset__

```{r}
df_party %>% summarise(`Minimum Year` = min(year(date_between)), `Maximum Year` = max(year(date_between)))
```

### Comparative Analysis

This section will compare the trajectory of INC and BJP at states since 1952.

#### Political Dominance {.tabset}

##### Count of CMs

```{r cm-count}
df_party %>%
  ggplot() +
  geom_line(aes(date_between, inc_states, color = "INC", group = 1), alpha = 0.5) +
  geom_smooth(aes(date_between, inc_states, color = "INC", group = 1), se = F, method = loess) +
  geom_line(aes(date_between, bjp_states, color = "BJP", group = 1), alpha = 0.5) +
  geom_smooth(aes(date_between, bjp_states, color = "BJP", group = 1), se = F, method = loess) +
  theme_bw() +
  scale_x_date(breaks = "5 years", date_labels = "%Y") +
  labs(x = "Year", y = "INC & BJP CMs (n)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

##### Percentage of CMs

```{r cm-perc}
df_party %>%
  ggplot() +
  geom_line(aes(date_between, inc_states*100/total_valid_states, color = "INC", group = 1), 
            alpha = 0.5) +
  geom_smooth(aes(date_between, inc_states*100/total_valid_states, color = "INC", group = 1), 
              se = F, method = loess) +
  geom_line(aes(date_between, bjp_states*100/total_valid_states, color = "BJP", group = 1), 
            alpha = 0.5) +
  geom_smooth(aes(date_between, bjp_states*100/total_valid_states, color = "BJP", group = 1), 
              se = F, method = loess) +
  theme_bw() +
  scale_x_date(breaks = "5 years", date_labels = "%Y") +
  scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  labs(x = "Year", y = "INC & BJP CMs (%)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

##### Population under CMs

```{r cm-pop}
df_party_pop %>%
  ggplot() +
  geom_line(aes(date_between, pop_inc, color = "INC", group = 1), 
            alpha = 0.5) +
  geom_line(aes(date_between, pop_bjp, color = "BJP", group = 1), 
              alpha = 0.5) +
  geom_smooth(aes(date_between, pop_inc, color = "INC", group = 1), 
              se = F, method = loess) +
  geom_smooth(aes(date_between, pop_bjp, color = "BJP", group = 1), 
              se = F, method = loess) +
  theme_bw() +
  scale_x_date(breaks = "5 years", date_labels = "%Y") +
  scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  labs(x = "Year", y = "Population under INC & BJP CMs (%)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

##### Mean Tenure

```{r cm-tenure}
df_party_tenure %>%
  ggplot() +
  geom_line(aes(date_between, tenure_inc, color = "INC", group = 1), 
            alpha = 0.5) +
  geom_line(aes(date_between, tenure_bjp, color = "BJP", group = 1), 
            alpha = 0.5) +
  geom_smooth(aes(date_between, tenure_inc, color = "INC", group = 1), 
              se = F, method = loess) +
  geom_smooth(aes(date_between, tenure_bjp, color = "BJP", group = 1), 
              se = F, method = loess) +
  theme_bw() +
  scale_x_date(breaks = "5 years", date_labels = "%Y") +
  labs(x = "Year", y = "Mean Tenure Duration of INC & BJP CMs") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

##### Tenure - Boxplots

```{r cm-tenure-bxplt}
df_final %>% 
  select(-date_between) %>% 
  distinct(state_name,name,party, .keep_all = T) %>%
  filter(party %in% c("INC", "BJP")) %>% 
  ggplot(aes(party, total_tenure, fill = party)) +
  geom_boxplot() +
  theme_bw() +
  labs(x = "Party", y = "Tenure Duration") +
  scale_fill_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top")
```

#### Social Background {.tabset}

__Simple Count of CMs__

```{r cm-soc-count-1}
df_final %>% 
  select(-date_between,) %>%
  distinct(unique_id, .keep_all = T) %>%
  filter(party %in% c("INC","BJP") & category_new != "PR") %>%
  count(party, category_new) %>% 
  mutate(category_new = factor(category_new,
                               levels = c("GEN","OBC","SC","ST","MUSLIM","CHRISTIAN","N/A"))) %>%
  arrange(category_new) %>% 
  pivot_wider(names_from = category_new, values_from = n, values_fill = 0)
```

__Unique Count of CMs__

```{r cm-soc-count-2}
df_final %>% 
  select(-date_between,) %>%
  distinct(name, .keep_all = T) %>%
  filter(party %in% c("INC","BJP") & category_new != "PR") %>%
  count(party, category_new) %>% 
  mutate(category_new = factor(category_new,
                               levels = c("GEN","OBC","SC","ST","MUSLIM","CHRISTIAN","N/A"))) %>%
  arrange(category_new) %>% 
  pivot_wider(names_from = category_new, values_from = n, values_fill = 0)
```

##### Simple Proportion

```{r cm-soc-prop-1}
df_final %>% 
  select(-date_between,) %>%
  distinct(unique_id, .keep_all = T) %>%
  filter(party %in% c("INC","BJP") & category_new != "PR") %>%
  count(party, category_new) %>% 
  mutate(category_new = factor(category_new,
                               levels = c("GEN","OBC","SC","ST","MUSLIM","CHRISTIAN","N/A"))) %>% 
  group_by(party) %>% 
  mutate(s_prop = n*100/sum(n)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_col(aes(category_new, s_prop, fill = party), position = "dodge") +
  theme_bw() +
  labs(x = "Social Identity", y = "Percentage (%)") +
  scale_fill_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top")
```

##### Unique Proportion

```{r cm-soc-prop2}
df_final %>% 
  select(-date_between,) %>%
  distinct(name, .keep_all = T) %>%
  filter(party %in% c("INC","BJP") & category_new != "PR") %>%
  count(party, category_new) %>% 
  mutate(category_new = factor(category_new,
                               levels = c("GEN","OBC","SC","ST","MUSLIM","CHRISTIAN","N/A"))) %>% 
  group_by(party) %>% 
  mutate(s_prop = n*100/sum(n)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_col(aes(category_new, s_prop, fill = party), position = "dodge") +
  theme_bw() +
  labs(x = "Social Identity", y = "Percentage (%)") +
  scale_fill_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top")
```

##### INC CMs Caste

```{r inc-soc}
df_party_cat %>% 
  ggplot() +
  geom_smooth(aes(date_between, gen_states_inc*100/inc_states, color = "GEN"), 
              se = F, method = loess) +
  geom_smooth(aes(date_between, obc_states_inc*100/inc_states, color = "OBC"),
              se = F, method = loess) +
  geom_smooth(aes(date_between, sc_states_inc*100/inc_states, color = "SC"), 
              se = F, method = loess) +
  geom_smooth(aes(date_between, st_states_inc*100/inc_states, color = "ST"), 
              se = F, method = loess) +
  theme_bw() +
  labs(x = "Year", y = "INC CMs - Caste Groups (%)") +
  scale_x_date(breaks = "5 years", date_labels = "%Y") +
  scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  scale_color_manual(values = c(GEN = "#1f77b4", OBC = "#ff7f0e",
                                SC = "#17becf", ST = "#2ca02c"),
                     labels = c(GEN = "GEN", OBC = "OBC", 
                                SC = "SC", ST = "ST"),
                     name = "Caste Group") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "top")
```

##### BJP CMs Caste

```{r bjp-soc}
df_party_cat %>% 
  ggplot() +
  geom_smooth(aes(date_between, gen_states_bjp*100/bjp_states, color = "GEN"), 
              se = F, method = loess) +
  geom_smooth(aes(date_between, obc_states_bjp*100/bjp_states, color = "OBC"),
              se = F, method = loess) +
  geom_smooth(aes(date_between, sc_states_bjp*100/bjp_states, color = "SC"), 
              se = F, method = loess) +
  geom_smooth(aes(date_between, st_states_bjp*100/bjp_states, color = "ST"), 
              se = F, method = loess) +
  theme_bw() +
  labs(x = "Year", y = "BJP CMs - Caste Groups (%)") +
  scale_x_date(breaks = "5 years", date_labels = "%Y") +
  scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  scale_color_manual(values = c(GEN = "#1f77b4", OBC = "#ff7f0e",
                                SC = "#17becf", ST = "#2ca02c"),
                     labels = c(GEN = "GEN", OBC = "OBC", 
                                SC = "SC", ST = "ST"),
                     name = "Caste Group") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "top")
```

## MP Analysis

### Comparative Analysis {.tabset}

#### Count of MPs

```{r mp-count}
df_mp %>% 
  filter(Party %in% c("INC","BJP") & Position == 1) %>% 
  count(Elec_year,Party) %>% 
  ggplot(aes(Elec_year, n, color = Party)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (n)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Percentage of MPs

```{r mp-perc}
df_mp %>%
  mutate(Party = case_when(Party == "INC" ~ "INC",
                           Party == "BJP" ~ "BJP",
                           T ~ "Others")) %>% 
  filter(Position == 1) %>% 
  count(Elec_year,Party) %>%
  group_by(Elec_year) %>% 
  mutate(n = round(n*100/sum(n))) %>% 
  filter(Party != "Others") %>% 
  ggplot(aes(Elec_year, n, color = Party)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (%)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Count of MPs (Unique)
```{r mp-unq-count}
df_mp %>% 
  filter(Party %in% c("INC","BJP") & Position == 1) %>% 
  distinct(Party,Candidate, .keep_all = T) %>% 
  count(Elec_year,Party) %>% 
  ggplot(aes(Elec_year, n, color = Party)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP Unique MPs (n)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

#### Percentage of Unique MPs

```{r mp-unq-perc}
df_mp %>%
  distinct(Party,Candidate, .keep_all = T) %>% 
  mutate(Party = case_when(Party == "INC" ~ "INC",
                           Party == "BJP" ~ "BJP",
                           T ~ "Others")) %>% 
  filter(Position == 1) %>% 
  count(Elec_year,Party) %>%
  group_by(Elec_year) %>% 
  mutate(n = round(n*100/sum(n))) %>% 
  filter(Party != "Others") %>% 
  ggplot(aes(Elec_year, n, color = Party)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (%)") +
  scale_color_manual(values = c(INC = "#1f77b4", BJP = "#ff7f0e"),
                     labels = c(INC = "INC", 
                                BJP = "BJP"),
                     name = "Party") +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Gender (n)

```{r mp-gender-count}
df_mp %>% 
  filter(Party %in% c("INC","BJP") & Position == 1) %>% 
  count(Elec_year,Party,Sex) %>% 
  ggplot(aes(Elec_year, n, color = Sex)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (n)") +
  scale_color_tableau() +
  facet_wrap(~ Party) +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Gender (%)

```{r}
df_mp %>% 
  filter(Party %in% c("INC","BJP") & Position == 1) %>% 
  count(Elec_year,Party,Sex)  %>%
  group_by(Elec_year,Party) %>% 
  mutate(n = round(n*100/sum(n))) %>% 
  ggplot(aes(Elec_year, n, color = Sex)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (%)") +
  scale_color_tableau() +
  facet_wrap(~ Party) +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Caste (n)

```{r}
df_mp %>% 
  filter(Party %in% c("INC","BJP") & Position == 1) %>% 
  count(Elec_year,Party,Constituency_Type) %>% 
  ggplot(aes(Elec_year, n, color = Constituency_Type)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (n)") +
  scale_color_tableau() +
  facet_wrap(~ Party) +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Caste (%)

```{r}
df_mp %>% 
  filter(Party %in% c("INC","BJP") & Position == 1) %>% 
  count(Elec_year,Party,Constituency_Type) %>%
  group_by(Elec_year,Party) %>% 
  mutate(n = round(n*100/sum(n))) %>% 
  ggplot(aes(Elec_year, n, color = Constituency_Type)) +
  geom_line(linewidth = 1) +
  theme_bw() +
  labs(x = "Year", y = "INC & BJP MPs (%)") +
  scale_color_tableau() +
  facet_wrap(~ Party) +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```